---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Required imports
```{r}
library(ggplot2)
library(dplyr)
```
## Loading and preprocessing the data  
Reading the csv file after unzipping and converting into Dates
```{r}
data <- read.csv(unzip("activity.zip"), na.strings = "NA", stringsAsFactors = FALSE)
str(data)
data$date <- as.Date(data$date, "%Y-%m-%d")
head(data,10)
```
## What is mean total number of steps taken per day?
### Total number of steps taken per day
```{r}
total_steps <- aggregate(steps~date, data, sum)
head(total_steps,10)
```
### Histogram of total steps taken per day
```{r}
ggplot(total_steps, aes(steps)) + geom_histogram(binwidth = 500) + labs(title="Histogram of Total Steps taken per day",x="Steps", y="Count")
```  

### Mean of Total steps
```{r}
mean(total_steps$steps, na.rm = TRUE)
```
### Median of Total steps
```{r}
median(total_steps$steps, na.rm = TRUE)
```
## What is the average daily activity pattern?
### Average of steps in intervals
```{r}
avg_steps_interval <- aggregate(steps~interval, data,mean)
ggplot(avg_steps_interval,aes(x=interval, y=steps))+geom_line() + labs(title="Average Daily Activity Pattern", x="Intervals", y="Number of Steps")
```

## Imputing missing values
### Summary of current dataset
```{r}
summary(data)
```  
From the summary it can be seen that 'steps' column has missing values
Replacing them with average of that particular interval in new dataset
```{r}
new_data <- data
head(new_data)
```
Getting indices of rows which have NA values
```{r}
na_index <- which(is.na(data$steps))
```
Replacing with average of that particular interval
```{r}
for (i in na_index){
        new_data[i,1] <- avg_steps_interval[which(avg_steps_interval$interval == new_data[i,3]),2]
}
## Checking if all NA values are replaced
summary(data)
summary(new_data)
```
### Differences between mean and median of both the datasets
```{r}
new_total_steps = aggregate(steps~date, new_data, sum)
mean(new_total_steps$steps)
median(new_total_steps$steps)
```  

## Are there differences in activity patterns between weekdays and weekends?  
### Creating new factor variables with 2 levels "Weekday" and "Weekend"
```{r}
new_data <- new_data %>% mutate(weekday = weekdays(data[,2]))
head(new_data)
tail(new_data)
```
Giving levels based on weekday
```{r}
new_data <- new_data %>% mutate(day_type = ifelse(weekday %in% c("Saturday", "Sunday"),"Weekend","Weekday"))
head(new_data)
tail(new_data)
```
Aggregating based on intervals and "Weekday" or "Weekend"
```{r}
avg_steps <- aggregate(steps~interval+day_type, new_data,mean)
head(avg_steps)
```
Plotting
```{r}
ggplot(avg_steps,aes(x=interval, y=steps)) + geom_line() + facet_grid(day_type ~ .) + labs(title="Average Steps in Intervals on Weekdays and Weekends",x="Intervals",y="Average Steps")
```
On weekdays average number of steps are higher than average number of steps on weekends